---
title: "p8105 HW1"
output: github_document
---
Author: "Jianyou Liu"

Date: 2018/09/16

# Problem 1

##The **code chunk** below creates a *data frame* consisting of:

1. A random sample of size 10 from a uniform [0, 5] distribution

2. A logical vector indicating whether elements of the sample are greater than 2

3. A length-10 character vector

4. A length-10 factor vector

```{r problem_1}
library(tidyverse)

set.seed(1)
unif <- runif(10, 0, 5)
unif[1:10]
mean(unif)

vec_logical <- c(unif)
vec_logical > 2
mean(vec_logical > 2)

vec_char <- vector("character", length = 10)
length(vec_char)
mean(vec_char)

vec_fac <- factor(c("male", "male", "male", "female", "male", "female", "male", "female", "male", "female"))
length(vec_fac)
mean(vec_fac)

```

The mean of the random sample is `r mean(unif)`

The mean of the logical vector is `r mean(vec_logical > 2)`

The mean of the character vector is `r mean(vec_char)`

The mean of the factor vector is `r mean(vec_fac)`

Only the first two variables could be taken means because they have numeric values assigned whereas the last two conprise only letters and characters.

## The **code chunk** below converts variables to numeric ones

```{r problem_1_conversion, results = 'hide'}
vec_logical_new <- as.numeric(vec_logical)
vec_char_new <- as.numeric(vec_char)
vec_fac_new <- as.numeric(vec_fac)
```
After conversion:
The mean of the random sample is `r mean(unif)`

The mean of the logical vector is `r mean(vec_logical_new > 2)`

The mean of the character vector is
`r mean(vec_char_new)`

The mean of the factor vector is `r mean(vec_fac_new)`

We are able to compute the mean for the factor vector but not for the character vector.

## The **code chunk** below converts variables to numeric ones with an *additional* step

```{r problem_1_conversion2, results = 'hide'}
vec_char_new2 <- as.numeric((as.factor(vec_char)))
vec_fac_new2 <- as.numeric((as.character(vec_fac)))
```
The mean of the factor vector is `r mean(vec_fac_new2)`

The mean of the character vector is
`r mean(vec_char_new2)`

We are now able to take the mean of the new character vector but not the factor vector anymore.


# Problem 2

##The **code chunk** below creates a *data frame* for problem2

```{r problem_2_df}
set.seed(123)

p2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_log2 = c(x + y > 0),
  vec_num2 = as.numeric(vec_log2),
  vec_fac2 = as.factor(vec_log2)
)

```
* The size of
  
  "x" is `r dim.data.frame(p2_df$x)`
  
  "y" is `r dim.data.frame(p2_df$y)`
  
  logical vector is `r dim.data.frame(p2_df$vec_log2)`
  
  numeric vector is `r dim.data.frame(p2_df$vec_num2)`
  
  factor vector is `r dim.data.frame(p2_df$vec_fac2)`

* The mean and median of "x" are `r mean(p2_df$x)` and `r median(p2_df$x)` repectively.

* The proportion of cases for which the logical vector is TRUE is equal to `r table(p2_df$vec_log2)["TRUE"] / length(p2_df$vec_log2)`

## The code chunk below displays scatterplots of variables "y" vs. "x" with different color points

```{r yx_scatter}
ggplot(p2_df, aes(x = x, y = y, color = vec_log2 )) + geom_point()

ggplot(p2_df, aes(x = x, y = y, color = vec_num2 )) + geom_point()

ggplot(p2_df, aes(x = x, y = y, color = vec_fac2 )) + geom_point()



```

